---
title: "Nonlinear with continuous predictor variables"
output: 
  html_notebook:
    toc: yes
---

# Load libraries
```{r}
library(olsrr)
library(psych)
library(tidyverse)
```



# Import data
```{r}
cog <- read_csv("cogtest.csv")
```

# describe data
```{r}
describe(cog)
```

# Plot the data
## As a linear relationship
```{r}
ggplot(cog, aes(x = practice, y = score)) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point() +
  scale_x_continuous(limits=c(0,14), breaks = seq(0, 14, by = 2)) +
  labs(title = "Does more practice equal better score?", subtitle = "Overlay best fit straight line", 
       x = "Minutes Spent Practicing", y = "Score")


```


## Plot data with a quadratic function
```{r}
# plot data with quadratic function
ggplot(cog, aes(x = practice, y = score)) + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE) + 
  geom_point() +
  scale_x_continuous(limits=c(0,14), breaks = seq(0, 14, by = 2)) +
  labs(title = "Does more practice equal better score?", subtitle = "Overlay quadratic curve", 
       x = "Minutes Spent Practicing", y = "Score")
```

# Mutate variables for polynomial regression: Get quadratic and cubic variables
```{r}
cog <- mutate(cog,
                  practice2 = practice^2,
                  practice3 = practice^3)
```



# Run linear, quadratic and cubic models
## Linear model
```{r}
mod_lin <- lm(score ~ practice, data = cog)
ols_regress(mod_lin)
```


## Quadratic model
```{r}
mod_quad <- lm(score ~ practice + practice2, data = cog)
ols_regress(mod_quad)
```

## Cubic model
```{r}
mod_cub <- lm(score ~ practice + practice2 + practice3, data = cog)
ols_regress(mod_cub)
```


# Assignment
# read in data
```{r}
hp <- read_csv("happyplanet.csv")
```


```{r}
hp <- hp %>% 
  mutate(gdp = gdp/10000)

write_csv(hp, "happyplanet2.csv", na = "")
```


# Describe data
```{r}
describe(hp)
```


# Plot the relationship between wellbe and gdp
```{r}
ggplot(hp, aes(x = gdp, y = wellbe)) +
  geom_point() +
  geom_smooth(method = "lm")
```




```{r}
ggplot(hp, aes(x = gdp, y = wellbe)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x,2), se = FALSE)
```

# Mutate the gdp variable
```{r}
hp <- mutate(hp, 
             gdp2 = gdp^2,
             gdp3 = gdp^3)
```



# Run models
## Linear
```{r}
mod_lin <- lm(wellbe ~ gdp, data = hp)
ols_regress(mod_lin)
```

## quadratic
```{r}
mod_quad <- lm(wellbe ~ gdp + gdp2, data = hp)
ols_regress(mod_quad)
```

## cubic
```{r}
mod_cub <- lm(wellbe ~ gdp + gdp2 + gdp3, data = hp)
ols_regress(mod_cub)
```



