---
title: "R Notebook"
output: html_notebook
---

# Load Libraries
```{r,message=FALSE}
library(tidyverse)
library(olsrr)
library(psych) 
```

# Read in dataset
```{r,message=FALSE}
slp <- read_csv("slpdata.csv")
```

# Reduce dataset to just variables of interest
```{r}
slp <- select(slp, cond, hygiene)
```



# Describe variables
```{r}
describe(slp)
```
categorical vars = cond, prior, & sex


# Part 1: Using dummy coding, use treatment condition, sex, and their interaction to predict sleep hygiene.

## Specify dummy codes for categorical predictors
```{r}

slp <- mutate(slp,
              cond2 = ifelse(cond == 2, 1, 0),
              cond3 = ifelse(cond == 3, 1, 0)) 
```

## Run model
```{r}
m1 <- lm(hygiene ~ cond2 + cond3, data = slp)
ols_regress(m1)
```
Interpretation of dummy coded model:

Intercept: The predicted sleep hygiene score when all x variables are zero, so participants in Condition 1.

cond2: the predicted difference in sleep hygiene score between particiants in condition 2 compared to condition 1.

cond3: the predicted difference in sleep hygiene score between particiants in condition 3 compared to condition 1.



# Part 2: Redo this analysis using effect coding for treatment condition

## Specificy effect coding for categotical variables
"With this coding structure the same process is completed as
that for dummy coding except the last group receives a −1 on all contrasts, thus
only k −1 contrasts are used in this coding type (Wendorf, 2004)." (from Davis, 2010)
```{r}
slp <- mutate(slp,
              
              cond2.ec = ifelse(cond == 2, 1, 0),
              cond3.ec = ifelse(cond == 3, 1, 0),
              
              cond2.ec = ifelse(cond == 1, (-1), cond2.ec),
              cond3.ec = ifelse(cond == 1, (-1), cond3.ec)) 
```

## Run model without interaction first
```{r}
m2 <- lm(hygiene ~ cond2.ec + cond3.ec, data = slp)
ols_regress(m2)
```
Interpretation of effect coding model:

Effects coding allows us to examine mean differences between two groups. In effects coding, the slope estimates are simply the difference between the mean of the group coded as 1 and the grand mean of all the groups.

Intercept: In effect coding, the intercept is the grand mean of sleep hygiene across all the three treatment groups

cond2: the predicted difference in sleep hygiene score between particiants in condition 2 compared to the mean of all three treatment conditions.

cond3: the predicted difference in sleep hygiene score between particiants in condition 3 compared to the mean of all three treatment conditions.


# Contrast Coding
```{r}

slp <- mutate(slp,
              contrast_1v2 = ifelse( cond == 1,  .5, ifelse(cond == 2, -.5,   0)),   # Compare condition 1 and condition 2
              contrast_3v12 = ifelse(cond == 1, -.5, ifelse(cond == 2, -.5,   1)))   # Compare condition 3 to condition 1 & 2
```

# Contrast Coded Model
```{r}
m3 <- lm(hygiene ~ contrast_1v2 + contrast_3v12, data = slp)
ols_regress(m3)
```


```{r}
aggregate(slp$hygiene, by = list(cond = slp$cond), mean)

 6.28730-1.627
 5.994+1.041
 
 
```











# Try it yoursefl activity
## Load libraries
```{r}
library(psych)
library(olsrr)
library(tidyverse)
```


## Read in data
```{r}

diet <- read_csv("Diet.csv")

```

# Reduce dataset
```{r}
diet_reduced <- select(diet, gender, Diet, weightLOST)
```


```{r}
describe(diet_reduced)

```

```{r}
diet_reduced %>% 
  group_by(Diet) %>% 
  summarize(mean_weightLOST = mean(weightLOST, na.rm = TRUE))
```


## Dummy code diet variable
```{r}
diet_reduced <- mutate(diet_reduced, 
                       Diet2.dum = ifelse(Diet == 2, 1, 0),
                       Diet3.dum = ifelse(Diet == 3, 1, 0)
                       )
```



## Run regression model with dummy coded variables
```{r}
mod1 <- lm(weightLOST ~ Diet2.dum + Diet3.dum, data = diet_reduced)
ols_regress(mod1)
```
```{r}
mod2 <- lm(weightLOST ~ gender + Diet2.dum + Diet3.dum + gender*Diet2.dum + gender*Diet3.dum, data = diet_reduced)
ols_regress(mod2)
```

## Effect coding 
```{r}

diet_reduced <- mutate(diet_reduced,
                       
                       Diet2.ec = ifelse(Diet == 2, 1, 0),
                       Diet3.ec = ifelse(Diet == 3, 1, 0),
                       
                       Diet2.ec = ifelse(Diet == 1, -1, Diet2.ec),
                       Diet3.ec = ifelse(Diet == 1, -1, Diet3.ec))


```


## Run regression model with effect coded variables
```{r}
mod3 <- lm(weightLOST ~ Diet2.ec + Diet3.ec, data = diet_reduced)
ols_regress(mod3)
```
```{r}
mod3 <- lm(weightLOST ~ Diet2.ec + Diet3.ec + gender*Diet2.ec + gender*Diet3.ec, data = diet_reduced)
ols_regress(mod3)
```
