---
title: 'PSY 653 Module 2: Interaction Effects (aka Moderation)'
author: "Neil Yetz & Gemma Wallace"
subtitle: Feb 5, 2020
output:
  pdf_document: default
  html_document:
    df_print: paged
---
# Part 1: In-class Demo
Use the "moderation_demo.csv" datafile

## Load libraries
```{r}
library(psych)
library(tidyverse)
library(olsrr)
```

## Read in data
```{r}
dat <- read_csv("moderation_demo.csv")
```

## Get descriptives
```{r}
describe(dat)
```

## Examine correlation between att1 & att3
```{r}
cor(dat$att1, dat$att3)
```
Since the correlation between att1 and att3 is small, there is room for a moderation effect.


```{r}
ggplot(dat, aes(x = att1, y = att3)) +
  geom_smooth(method = "lm")+
  geom_jitter()+
  theme_classic()
```



## Create the cross product of att1 & att3
```{r}
dat <- mutate(dat, att1att3 = att1*att3)
```

## Run moderated regression in which att1, att3, and att1*att3 predict out4
### Main Effects model
```{r}
modME <- lm(out4  ~ att1 + att3, data = dat)
ols_regress(modME)
```

### Interaction model
```{r}
modINT <- lm(out4 ~ att1 + att3 + att1*att3 , data = dat)
ols_regress(modINT)
```


```{r}



library(sjPlot)
library(sjmisc)


plot_model(modINT, type = "pred", terms = c("att3", "att1"),
           show.intercept = TRUE, se = FALSE)


```




# Compare the two models
```{r}
anova(modME,modINT)
```
The model that includes the interaction effect accounts (modINT) for a significantly larger amount of variance in out4 compared to the model that only includes the main effects (modME) of the two predictors.

# Part 2: Try it yourself
Use the "moderation_sleepdata.csv" datafile
In this exmaple, you will examine how sleep hygiene, anxiety, and the interaction between sleep hygiene and anxiety influence how well participants slept.

## Read in data
```{r}
sleep_data <- read_csv("moderation_sleepdata.csv")
```

## Get descriptives
```{r}
describe(sleep_data)
```

## Examine correlation between anxiety and hygiene
```{r}
cor(sleep_data$anxiety, sleep_data$hygiene)
```
Since the correlation between the two variables is relatively small, there is room for a moderation effect

## Create the cross-product of hygiene and anxiety
```{r}
sleep_data <- mutate(sleep_data, anxiety_hygiene = hygiene*anxiety)
```

## Run moderated regression in which hygiene, anxiety, and hygiene*anxiety predict sleep
### Main effects model
```{r}
slp_mod1 <- lm(sleep  ~ hygiene + anxiety, data = sleep_data )
ols_regress(slp_mod1)
```

### Interaction model
```{r}
slp_mod2 <- lm(sleep  ~ hygiene + anxiety + anxiety_hygiene, data = sleep_data )
ols_regress(slp_mod2)
```

## Compare the two models
```{r}
anova(slp_mod1,slp_mod2)
```
At a significance threshold of p<0.05, the model that included the interaction effect (slp_mod2) explained a significantly larger amount of variance in sleep than the model that only tested the main effects of anxiety and sleep hygiene (slp_mod1).
