---
title: "R Notebook"
output: html_notebook
---

# Kevin's R code
```{r}
dfhyp= NA				
dferr= NA
alpha=.05
effect=	NA				#enter the minimum effect you are testing
sse=100	        #Don't change this				
mse=(((1/effect)-1)*sse)/dferr
noncen=sse/mse
qf((1-alpha),dfhyp,dferr,noncen)
```


# 1a Applied to Bret et al. (Right Wing Authoritarianism)
```{r}
dfhyp=1					
dferr=161
alpha=.05
effect=.01					#enter the minimum effect you are testing
sse=100					
mse=(((1/effect)-1)*sse)/dferr
noncen=sse/mse
qf((1-alpha),dfhyp,dferr,noncen)

```

The authors need an F-value of 8.683388 to reach a significant effect at a minimum effect size of .01.

# 1b
With their F-value of 4.81. They had enough power to reject the comparison of the nil (F-value = 3.89)
With their F-value of 4.81. They did not have enough power to reject the comparison of 1% or less of the variance (F = 8.61)

# 1c
They would need a sample size of at least 122 (dferror = n - k) to have a power of .80.
They would need a sample size of at least 302 to reject the 1% or less of the variance and have a power of .80. 



# Load libraries
```{r}
library(tidyverse)
library(psych)
library(car)
library(BayesFactor)
```

# read in retirement.csv
```{r}
retirement <- read_csv("retirement.csv")
```

# Factor the categorical variables
```{r}
retirement <- mutate(retirement,
                     sex.f = factor(sex, 
                                       levels = c(1,2), 
                                       labels = c("Female", "Male")),
                     occupation.f = factor(occupation, 
                                           levels = c(1,2,3,4,5,6), 
                                        labels = c("Prof", "Manag", "nonmanual", "Skilled", "Semi-Skilled", "Unskilled")))
```




# describe data
```{r}
describe(retirement)
```


# Normal ANOVA
```{r}
anova(lm(mental~ sex.f*occupation.f, data = retirement))
```
```{r}
sex_eta    <-  0.21 / (1529.99 + .21 + 45.93 + 6.01)
occ_eta    <- 45.93 / (1529.99 + .21 + 45.93 + 6.01)
int_et     <-  6.01 / (1529.99 + .21 + 45.93 + 6.01)

sex_eta
occ_eta   
int_et    
```



# Bayes ANOVA: "all"
```{r}

m1 <- anovaBF(mental~ sex.f*occupation.f, data = retirement, whichModels = "all")
m1
```
```{r}
plot(m1)
```


#Bayes ANOVA: "top"
```{r}
m2 <- anovaBF(mental~ sex.f*occupation.f, data = retirement, whichModels = "top")
m2
```


```{r}
plot(m2)
```


#Bayes ANOVA: "bottom"
```{r}
m3 <- anovaBF(mental~ sex.f*occupation.f, data = retirement, whichModels = "bottom")
m3
```

```{r}
plot(m3)
```




# Try it yourself


## 1a Applied to Bret et al. (Right Wing Authoritarianism)
```{r}
dfhyp=1					
dferr=230
alpha=.05
effect=.01					#enter the minimum effect you are testing
sse=100					
mse=(((1/effect)-1)*sse)/dferr
noncen=sse/mse
qf((1-alpha),dfhyp,dferr,noncen)

```

## 1b
compared to nill = 3.88
compared to 1% = 9.58


## 1c
At Nil the needed sample size is 71 (dferror = n - 1)
At 1% The needed samples size is 121 (dferror = n - 1)


## Part 2
```{r}

slp <- read_csv("slpdata.csv")

```

```{r}
slp <- mutate(slp, 
             sex.f = factor(sex, levels = c(1,2), labels = c("male", "female")),
             cond.f = factor(cond,  levels = c(1,2,3), labels= c("selfhelp", "group", "group_and_partner"))
             ) 
```



```{r}
anov <- lm(sleep ~ sex.f*cond.f, data=slp)
anova(anov)

```


```{r}
Bayesm1 <- anovaBF(sleep ~ sex.f*cond.f, data=slp,whichModels = 'all')
Bayesm1
plot(Bayesm1)

```


```{r}
Bayesm2 <- anovaBF(sleep ~ sex.f*cond.f, data=slp, whichModels = "top")
Bayesm2
plot(Bayesm2)
```



```{r}
Bayesm3 <- anovaBF(sleep ~ sex.f*cond.f, data=slp,whichModels = "bottom")
Bayesm3
plot(Bayesm3)

```





