---
title: "R Notebook"
author: "Neil Yetz & Gemma Wallace"
output:
  pdf_document: default
  html_notebook: default
---

# Load Libraries
```{r}
library(psych)
library(tidyverse)
library(lme4)
library(lmerTest)
library(broom)
library(modelr)
```


# class demo dataset
```{r}
#Wide dataset will convert to long below
wide <- read_csv("mlm_grow.csv")
```

# Format and output class demo
```{r}
###### Stolen from Kim Start #######
long_perf <- wide %>%
  select(kid_id, txcond, perf_0:perf_6) %>%
  gather(key = measure, value = perform, perf_0:perf_6) %>%
  separate(measure, c("temp", "week")) %>%
  select(-temp) %>%
  mutate(week = as.numeric(week))


long_inter <- wide %>%
  select(kid_id, inter_0:inter_6) %>%
  gather(key = measure, value = interest, inter_0:inter_6) %>%
  separate(measure, c("temp", "week")) %>%
  select(-temp) %>%
  mutate(week = as.numeric(week))

grow <- full_join(long_perf, long_inter, by = c("kid_id", "week"))

grow <- grow %>%
  mutate(txcond.f = factor(txcond, levels = c(0,1), labels = c("control", "treatment")),
  kid_id.f = factor(kid_id))

###### Stolen From Kim End ######

# Simplify dataset
grow <- select(grow, kid_id, week, perform)
# Write to csv
#write_csv(grow, "grow.csv", na = "")

```

# Start demo

# Read in Data
```{r}
grow <- read_csv("grow.csv")
```


# Describe data
```{r}
describe(grow)
```

# Visualize the data
```{r, fig.width=9, fig.height=9}

ggplot(grow, aes(x = week, y = perform)) +
  geom_jitter(aes(color = factor(week))) +
  geom_smooth(method = "loess", color = "red", size = 2) +
  xlab("Week") +
  ylab("Performance") +
  labs(color = "Week")
```

# Baseline model
```{r}
mod1 <-  lmer(perform ~ 1 + (1|kid_id), REML = TRUE, data = grow)
summary(mod1)
```






# Separated by subject
```{r, fig.height = 12, fig.width=12}
ggplot(data = grow, aes(x = week, y = perform)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_continuous(limits = c(0,10)) +
  scale_x_continuous(limits = c(0,6), breaks = c(0,1,2,3,4,5,6)) +
  facet_wrap(~kid_id) +
  labs(title = "Do students in the control condition improve their task performance over the course of the program?",
    x = "Week in the program", y = "Task Performance")
```




```{r}
mod2 <-  lmer(perform ~ 1 + week + (1 + week|kid_id), REML = TRUE, data = grow)
summary(mod2)
```

```{r, fig.width=12, fig.height=8}
# Get predicted values
mod2.plot <- add_predictions(data = grow, model = mod2)

# Make plot
ggplot(data = mod2.plot, aes(x = week, y = pred, group = kid_id)) +
  geom_line(color = "grey53") +
  geom_abline(intercept = 4.4731, slope = .1438, color="red", size=3) +
  scale_y_continuous(limits = c(0,10)) +
  scale_x_continuous(limits = c(0,6), breaks = c(0,1,2,3,4,5,6)) +
  labs(title = "Do students improve on task performance over the course of the program?",
    x = "Week", y = "Predicted Performance") +
  theme_bw()


```




# Assignment
# read in data
```{r}
long <- read_csv("Longitudinal.csv")
```

# Describe data
```{r}
describe(long)
```


#Factor variables
```{r}
long <- mutate(long, 
               job = factor(job, levels = c(0,1), labels = c("half time", "full time")),
               sex = factor(sex, levels = c(0,1), labels = c("male", "female"        ))
               )
```

# Create occas squared
```{r}
long <- mutate(long, 
               occassq = occas**2)
```



# Aggregate data
```{r}
agg_long <- aggregate(x=long$gpa,by=list(occasion = long$occas), FUN=mean)
agg_long
```



```{r}
ggplot(agg_long, aes(x = occasion, y = x)) +
  geom_line() +
  ylab("GPA")
  
```




```{r, fig.width=9, fig.height=9}

ggplot(long, aes(x = occas, y = gpa)) +
  geom_jitter(aes(color = factor(occas))) +
  geom_smooth(method = "loess", color = "red", size = 2) +
  xlab("Occasion") +
  labs(color = "Occasion")
```

```{r, fig.width = 9, fig.height=9}

ggplot(long, aes(x = occas, y = gpa)) +
  geom_jitter(aes(color = factor(occas))) +
  geom_smooth(method = "loess", color = "red", size = 2) +
  xlab("Occasion") +
  labs(color = "Occasion") +
  facet_wrap(~sex)
```



```{r, fig.width = 12, fig.height=12}
ggplot(data = long, aes(x = occas, y = gpa)) +
geom_point() +
geom_smooth(method = "lm", se = FALSE) +
facet_wrap(~student) +
labs(title = "Do students increase in GPA across occasion?",
x = "Occasion", y = "GPA")
```


# Baseline model
```{r}

mod1 = lmer(gpa ~ 1 + (1|student), REML = TRUE, data = long)
summary(mod1)
```



# Added time as a random factor
```{r}

mod2 = lmer(gpa ~ 1 + occas + (1 + occas|student), REML = TRUE, data = long)
summary(mod2)
```

# ICC
ICC = variance for effect of interest / variance for effect + unexplained variance
This is the underlying structure of the F statistic
ICC for mod 2 = .005 / .005 + .042 = .106

Compare the ICCs across the different models 



# Add covariate
```{r}

mod4 = lmer(gpa ~ 1 + sex + (1|student), REML = TRUE, data = long)
summary(mod4)
```


```{r}
mod5 = lmer(gpa ~ 1 + occas + sex + (1 + occas|student), REML = TRUE, data = long)
summary(mod5)
```

# Job + sex
```{r}
mod6 = lmer(gpa ~ 1 + sex + job + (1|student), REML = TRUE, data = long)
summary(mod6)
```


```{r}
mod7 = lmer(gpa ~ 1 + occas + sex + job + (1+ occas|student), REML = TRUE, data = long)
summary(mod7)
```