---
title: "R Notebook"
output: html_notebook
---

# Load libraries
```{r,message=FALSE}
library(tidyverse)
library(psych)
library(mediation)
library(ppcor)
```

# Read in data
```{r}
med <- read_csv("mediate2.csv")
```

# Class Demo
## Baron & Kenny Plausibility: Examine Correlations between variables
## Analysis 1: Test the hypothesis that X4 mediates the relationship between X1 and Y1

```{r}
cor(med)
```

r xy = 0.3465 (c path)
r xm = 0.043 (a path)
r my = 0.4105 (b path)

We do not have justification to test this mdiation hypothesis because the a path has a negligible effect.

## Analysis 2: Test the hypothesis that X4 mediates the relationship between X3 and Y1

r xy = 0.5053
r xm = 0.3425
r my = 0.4105

We do have justifaction to test this mediation hypothesis because all paths have a moderate effect

```{r}
spcor.test(x = med$X3, y = med$Y1, z = med$X4)
```

r y(x.m) = 0.3999. This is 0.11 smaller than r xy (0.5053), indicating that partial mediation is plausible. In other words, there is a portion of the relation betweeen x and y that involves m.


### Regression method
```{r}
m1 <- lm(Y1 ~ X3      , data = med)
m2 <- lm(Y1 ~ X3 + X4 , data = med)

anova(m1,m2)

```



### mediate in psych
```{r}

fitmed <- psych::mediate(Y1 ~ X3 + (X4), data = med)
summary(fitmed)

```



### Mediate in mediation package
```{r}
fitM <- lm(X4 ~ X3, data = med)
fitY <- lm(Y1 ~ X3 + X4, data = med)


fitmed <- mediation::mediate(fitM, fitY, treat = "X3", mediator = "X4")
summary(fitmed)

```

```{r}
cor(med)
```

## Analysis 3: Test the hypothesis that X4 mediates the relationship between X5 and Y1

r xy = 0.6405
r xm = 0.4068
r my = 0.4105

We do have justifaction to test this mediation hypothesis because all paths have a moderate effect

```{r}
spcor.test(x = med$X5, y = med$Y1, z = med$X4)
```


r y(x.m) = 0.5192. This is 0.12 smaller than r xy (0.6405), indicating that partial mediation is plausible. In other words, there is a portion of the relation betweeen x and y that involves m.



### Regression method
```{r}
m1 <- lm(Y1 ~ X5      , data = med)
m2 <- lm(Y1 ~ X5 + X4 , data = med)

anova(m1,m2)

```


### mediate in psych
```{r}

fitmed <- psych::mediate(Y1 ~ X5 + (X4), data = med)
summary(fitmed)

```


### Mediate in mediation package
```{r}
fitM <- lm(X4 ~ X5, data = med)
fitY <- lm(Y1 ~ X5 + X4, data = med)


fitmed <- mediation::mediate(fitM, fitY, treat = "X5", mediator = "X4")
summary(fitmed)

```


# Class Activity

## Read data
```{r}
slp <- read_csv("slpdata.csv")
```



## Baron & Kenny Plausibility: Examine Correlations between variables
```{r}
cor(slp)
```

## Analysis 1: Test the hypothesis that sleep efficiency mediates the relationship between sleep hygiene and life satisfaction.

r xy = 0.5053
r xm = 0.7332
r my = 0.6405

We do have justifaction to test this mediation hypothesis because all paths have a moderate to large effect

```{r}
spcor.test(x = slp$hygiene, y = slp$lifesat, z = slp$sleep)
```



r y(x.m) = 0.0464. This is 0.12 smaller than r xy (0.5053), indicating that partial mediation is plausible. In other words, there is a portion of the relation betweeen x and y that involves m.


### Regression method
```{r}
m1 <- lm(lifesat ~ hygiene         , data = slp)
m2 <- lm(lifesat ~ hygiene + sleep , data = slp)

anova(m1,m2)
```

### mediate in psych
```{r}

fitmed <- psych::mediate(lifesat ~ hygiene + (sleep), data = slp)
summary(fitmed)

```


### Mediate in mediation package
```{r}
fitM <- lm(sleep   ~ hygiene        , data = slp)
fitY <- lm(lifesat ~ hygiene + sleep, data = slp)


fitmed <- mediation::mediate(fitM, fitY, treat = "hygiene", mediator = "sleep")
summary(fitmed)

```

## Analysis 2: Test the hypothesis that sleep efficiency mediates the relationship between anxiety and life satisfaction.
### Baron & Kenny Plausibility: Examine Correlations between variables
```{r}
cor(slp)
```

r xy = -0.3228
r xm = -0.1310
r my = 0.6405

We do NOT have justifaction to test this mediation hypothesis because path a has a negligible correlation effect size (-.1310)


```{r}
spcor.test(x = slp$anxiety, y = slp$lifesat, z = slp$sleep)
```

r y(x.m) = -0.3110. This is 0.0118 only slighlty larger (closer to zero) than r xy (-0.3228), indicating that partial mediation is NOT plausible.



