---
title: "R Notebook"
output: html_notebook
---

# Kevin's R code
```{r}
dfhyp= NA				
dferr= NA
alpha=.05
effect=	NA				#enter the minimum effect you are testing
sse=100	        #Don't change this				
mse=(((1/effect)-1)*sse)/dferr
noncen=sse/mse
qf((1-alpha),dfhyp,dferr,noncen)
```


# 1a Applied to Bret et al. (Right Wing Authoritarianism)
```{r}
dfhyp=1					
dferr=161
alpha=.05
effect=.01					#enter the minimum effect you are testing
sse=100					
mse=(((1/effect)-1)*sse)/dferr
noncen=sse/mse
qf((1-alpha),dfhyp,dferr,noncen)

```

The authors need an F-value of 8.683388 to reach a significant effect at a minimum effect size of .01.

# 1b
With their F-value of 4.81. They had enough power to reject the comparison of the nil (F-value = 3.89)
With their F-value of 4.81. They did not have enough power to reject the comparison of 1% or less of the variance (F = 8.61)

# 1c
They would need a sample size of at least 122 (dferror = n - k) to have a power of .80.
They would need a sample size of at least 302 to reject the 1% or less of the variance and have a power of .80. 



# Load libraries
```{r}
library(tidyverse)
library(BayesFactor)
```

# read in retirement.csv
```{r}
retirement <- read_csv("retirement.csv")
```

# Factor the categorical variables
```{r}
retirement <- mutate(retirement,
                     sex.f = factor(sex, 
                                       levels = c(1,2), 
                                       labels = c("Female", "Male")),
                     occupation.f = factor(occupation, 
                                           levels = c(1,2,3,4,5,6), 
                                           labels = c("Prof", "Manag", "nonmanual", "Skilled", "Semi-Skilled", "Unskilled")))
```

# Frequentist  ANOVA
```{r}
summary(aov(lm(mental~ sex.f*occupation.f, data = retirement)))
```
```{r}
sex_eta    <-  0.2 / (1530 + .2 + 45.9 + 6)
occ_eta    <- 45.9 / (1530 + .2 + 45.9 + 6)
int_et     <-  6.0 / (1530 + .2 + 45.9 + 6)

sex_eta
occ_eta   
int_et    
```



# Bayes ANOVA: "all"
```{r}

m1 <- anovaBF(mental~ sex.f*occupation.f, data = retirement, whichModels = "all")
m1
plot(m1)

```
#Bayes ANOVA: "top"
```{r}

m2 <- anovaBF(mental~ sex.f*occupation.f, data = retirement, whichModels = "top")
m2
plot(m2)
```

#Bayes ANOVA: "bottom"
```{r}
m3 <- anovaBF(mental~ sex.f*occupation.f, data = retirement, whichModels = "bottom")
m3
plot(m3)
```





# Try it yourself
## 1a 
```{r}
dfhyp=1					
dferr=230
alpha=.05
effect=.01					#enter the minimum effect you are testing
sse=100					
mse=(((1/effect)-1)*sse)/dferr
noncen=sse/mse
qf((1-alpha),dfhyp,dferr,noncen)
```

## 1b
compared to nill = 3.88
compared to 1% = 9.58

## 1c
At Nil the needed sample size is 71 (dferror = n - 1)
At 1% The needed samples size is 121 (dferror = n - 1)


## Part 2
```{r,message=FALSE}
slp <- read_csv("slpdata.csv")
```

### Factor categorical predictors
```{r}
slp <- mutate(slp, 
             sex.f = factor(sex, levels = c(1,2), labels = c("male", "female")),
             cond.f = factor(cond,  levels = c(1,2,3), labels= c("selfhelp", "group", "group_and_partner"))
             ) 
```

### Frequentist ANOVA
```{r}
anova <- aov(sleep ~ sex.f*cond.f, data=slp)
summary(anova)
```

#### Eta squared calculations
```{r}
sex   <-  16182/ (46531 + 16182 + 24971 + 593)
condition    <- 24971 / (46531 + 16182 + 24971 + 593)
interaction     <-  593 / (46531 + 16182 + 24971 + 593)

sex
condition
interaction
```

## Bayesian ANOVA with all models
```{r}
Bayesm1 <- anovaBF(sleep ~ sex.f*cond.f, data=slp,whichModels = 'all')
Bayesm1
plot(Bayesm1)
```


## Bayesian ANOVA top-down approach
```{r}
Bayesm2 <- anovaBF(sleep ~ sex.f*cond.f, data=slp, whichModels = "top")
Bayesm2
plot(Bayesm2)
```


## Bayesian ANOVA bottom-up approch
```{r}
Bayesm3 <- anovaBF(sleep ~ sex.f*cond.f, data=slp,whichModels = "bottom")
Bayesm3
plot(Bayesm3)

```





