---
title: "R Notebook"
output: html_notebook
---

# Clear environment
```{r}
rm(list=ls())
```

# Load Libraries
```{r}
install.packages("MuMIn")
install.packages("modelr")

library(tidyverse)
library(lme4)
library(lmerTest)
library(psych)
library(MuMIn)
library(modelr)

```

# Read in Kim's data: Demo activity
```{r}
teams <- read_csv("mlm_teams.csv")
```

## Describe
```{r}
describe(teams)
```

## Factor txcond
```{r}
teams <- teams %>%
mutate(txcond.f = factor(txcond, levels = c(0,1), labels = c("control", "treatment")),
       team_id.f = factor(team_id))
```


## Factor team_id
```{r}
teams <- mutate(teams, team_id.f = factor(team_id))
```

# Get Mean of Means
```{r}

team_means  <- group_by(teams, team_id)
team_means  <- summarize(team_means, mean_score = mean(score))
meanofmeans <- summarize(team_means, meanofmeans = mean(mean_score))
meanofmeans

```

```{r}

ggplot(data = teams, aes(x = team_id.f, y = score)) +
  geom_boxplot() +
  stat_summary(aes(y = score, group = team_id.f), fun = mean, color = "red", geom = "point", pch = 17, size = 1.5) +
  geom_hline(yintercept = meanofmeans$meanofmeans, linetype="dashed", color = "black", size = 1) +
  scale_y_continuous(limits = c(20,100), breaks = seq(20, 100, 10)) +
  labs(title = "Mean and variability of final scores across teams",
       subtitle = "dashed line = mean of team means, red triangle = team mean",
       x = "Team ID", y = "Final Score") +
  theme(axis.text.x = element_text(colour="grey20", size=8, angle=90, hjust=.5))
```




## Models

```{r}
# Random intercept
randint <- lmer(score ~ 1 + (1|team_id.f), data = teams, REML = FALSE)
summary(randint)

```

```{r}
(80.62 / (80.62 + 71.48))

library(sjstats)
performance::icc(randint)

```

```{r}
r.squaredGLMM(randint)
```

```{r}
# Random intercept Fixed slope
rifs<-lmer(score ~ 1 + risk + (1|team_id.f),data=teams, REML = FALSE)
summary(rifs)
```

```{r}
performance::icc(rifs)

83.33 / (83.33 + 57.96)


```

```{r}
r.squaredGLMM(rifs)
```


```{r}
# random slopes fixed intercepts
rsfi<- lmer(score ~ 1 + risk + (0 + risk|team_id.f),data=teams, REML = FALSE)
summary(rsfi)
```

```{r}
performance::icc(rsfi)

11.32 / (11.32 + 73.37)


```

```{r}
r.squaredGLMM(rsfi)
```


```{r}
# all random
allrand<- lmer(score ~ risk +  (1 + risk|team_id.f),data=teams, REML = FALSE)
summary(allrand)
```
```{r}
performance::icc(allrand)

(79.603 + 7.213) / (79.603 + 7.213 + 40.206) 
```



```{r}
library(modelr)

# Model without with only random intercept
mod2_b.plot <- add_predictions(data = teams, model = randint)

ggplot(data = mod2_b.plot, aes(x = risk, y = pred, group = team_id.f)) +
  geom_line() +
  geom_abline(intercept = 77.868, slope = -2.326, color="red", size=3) +
  labs(title = "Do students with a higher risk index perform worse on the final test?", x = "Risk Quintile", y =
         "Predicted Final Score")

# Model without random slope
mod2_b.plot <- add_predictions(data = teams, model = rifs)

ggplot(data = mod2_b.plot, aes(x = risk, y = pred, group = team_id.f)) +
  geom_line() +
  geom_abline(intercept = 77.868, slope = -2.326, color="red", size=3) +
  labs(title = "Do students with a higher risk index perform worse on the final test?", x = "Risk Quintile", y =
         "Predicted Final Score")


# Model without random intercept
mod2_b.plot <- add_predictions(data = teams, model = rsfi)

ggplot(data = mod2_b.plot, aes(x = risk, y = pred, group = team_id.f)) +
  geom_line() +
  geom_abline(intercept = 77.868, slope = -2.326, color="red", size=3) +
  labs(title = "Do students with a higher risk index perform worse on the final test?", x = "Risk Quintile", y =
         "Predicted Final Score")

# Model with random slope & Intercept
mod2_a.plot <- add_predictions(data = teams, model = allrand)

ggplot(data = mod2_a.plot, aes(x = risk, y = pred, group = team_id.f)) +
  geom_line() +
  geom_abline(intercept = 77.868, slope = -2.326, color="red", size=3) +
  labs(title = "Do students with a higher risk index perform worse on the final test?", x = "Risk Quintile", y = "Predicted Final Score")
```




```{r}
r.squaredGLMM(allrand)
```


## Pairwise comparison
```{r}

anova(randint,allrand)

```
```{r}
anova(rifs,allrand)
```

```{r}
anova(rsfi,allrand)
```





# New Dataset: Try it Yourself
## Read in data
```{r}
#library(haven)
#
#test <- read_sav("popular2.sav")
#
#
#popular <- test %>% 
#  select(pupil, class, Cextrav, popular) %>% 
#  as_tibble()
#
#write_csv(popular, "popular.csv", na = "")

popular <- read_csv("popular.csv")


```


## Describe data
```{r}
describe(popular)
```

# Factor class
```{r}
popular <- mutate(popular, class.f = factor(class))
```

# Get Mean of Means
```{r}
team_means  <- group_by(popular, class)
team_means  <- summarize(team_means, mean_pop = mean(popular))
meanofmeans <- summarize(team_means, meanofmeans = mean(mean_pop))
meanofmeans
```

```{r}
ggplot(data = popular, aes(x = class.f, y = popular)) +geom_boxplot() +stat_summary(aes(y = popular, group = class.f), fun = mean, color = "red", geom = "point", pch = 17, size = 1.5) +geom_hline(yintercept = meanofmeans$meanofmeans, linetype="dashed", color = "black", size = 1) +scale_y_continuous(limits = c(0, 10), breaks = seq(0, 10, 0.5)) +labs(title = "Mean and variability of popularity across class",subtitle = "dashed line = mean of class means, red triangle = class mean",x = "Class", y = "Popularity") +theme(axis.text.x = element_text(colour="grey20", size=8, angle=90, hjust=.5))
```


## Random Intercept only model
```{r}
# Random intercept
randint <- lmer(popular ~ 1 + (1|class.f), data = popular, REML = FALSE)
summary(randint)
```

### ICC 
```{r}
# ICC
0.6945/(0.6945+1.2218)
```

### Pseudo R^2
```{r}
# R^2
r.squaredGLMM(randint)
```



## Random intercept Fixed Slope model
```{r}
# Random intercept Fixed slope
rifs<-lmer(popular ~ 1 + Cextrav + (1|class.f),data=popular, REML = FALSE)
summary(rifs)
```


### Plot
```{r}
mod2.plot <- add_predictions(data=popular, model=rifs)

ggplot(data=mod2.plot, aes(x=Cextrav, y=pred, group=class.f))+
  geom_line()+
  geom_abline(intercept=5.07824, slope=0.48620, color="red", size=3)+
  labs(title="Are extraverted students more popular?", x="Extraversion (Mean centered)", y ="Predicted Popularity")


```



### ICC
```{r}
# ICC
0.8314/(0.8314 + 0.9299)
```


### Pseudo R^2
```{r}
# R^2
r.squaredGLMM(rifs)
```



## Random Slope, Fixed Intercept model
```{r}

# random slopes fixed intercepts
rsfi<- lmer(popular ~ 1 + Cextrav + (1 + Cextrav|class.f),data=popular, REML = FALSE)
summary(rsfi)

```


### Plot model
```{r}
# Model without random intercept
mod3.plot <- add_predictions(data=popular, model=rsfi)

ggplot(data=mod3.plot, aes(x=Cextrav, y=pred, group=class.f))+
  geom_line()+
  geom_abline(intercept=5.06055, slope=0.34167, color="red", size=3)+
  labs(title="Are extraverted students more popular?", x="Extraversion (Mean Centered)", y ="Predicted Popularity")


```

### ICC
```{r}
#ICC
0.08356/(0.08356 + 1.59167)
```

### Pseudo R^2
```{r}
#R^2
r.squaredGLMM(rsfi)
```


## All Random Model
```{r}

# all random
allrand<- lmer(popular ~ Cextrav +  (1 + Cextrav|class.f),data=popular, REML = FALSE)
summary(allrand)

```

## Plot
```{r}
mod4.plot <- add_predictions(data=popular, model=allrand)

ggplot(data=mod4.plot, aes(x=Cextrav, y=pred, group=class.f))+
  geom_line()+
  geom_abline(intercept=5.03184, slope=0.49290, color="red", size=3)+
  labs(title="Are extraverted students more popular?", x="Extraversion (Mean Centered)", y ="Predicted Popularity")
```

### ICC
```{r}
#ICC
0.88225+0.02522/(0.88225+0.02522+0.89505)
```

### Psuedo R^2
```{r}
#R^2
r.squaredGLMM(allrand)
```


## Pairwise comparison
### randint vs all random
```{r}
anova(randint,allrand)
```
### RIFS vs all random
```{r}
anova(rifs,allrand)
```
### RSFI vs all random
```{r}
anova(rsfi,allrand)
```





# OLD DATA

```{r}
org <- read_csv("orgdata.csv")
```


## Describe
```{r, rows.print = 13}
describe(org)
```


## Create measure of gender balance in workforce
```{r}
org <- mutate(org, percmale = nmale/nemploy)
```


##  Models
```{r}
# Random intercept
randint <- lmer(Develop ~ 1 + (1|HCollect), data = org, REML = FALSE)
summary(randint)

# Random intercept Fixed slope
rifs<-lmer(Develop~ 1 + percmale + (1|HCollect),data=org, REML = FALSE)
summary(rifs)

# random slopes fixed intercepts
rsfi<- lmer(Develop ~ 1 + percmale + (0 + percmale|HCollect),data=org, REML = FALSE)
summary(rsfi)

# all random
allrand<- lmer(Develop ~ percmale +   (1 + percmale|HCollect),data=org, REML = FALSE)
summary(allrand)
```



## ICC
variance / (variance + residual)
```{r}
library(sjstats)

performance::icc(randint)
performance::icc(rifs)
performance::icc(rsfi)
performance::icc(allrand)

```


## R^2 
```{r, warning = FALSE}

r.squaredGLMM(randint)
r.squaredGLMM(rifs)
r.squaredGLMM(rsfi)
r.squaredGLMM(allrand)

```


## Pairwise comparison
```{r}
anova(rifs,allrand)
anova(randint,allrand)
anova(rsfi,allrand)
```
