---
title: 'PSY 653 Module 3: Nested Designs'
author: "Gemma Wallace and Neil Yetz"
date: "February 12, 2020"
output:
  html_document:
    df_print: paged
---

# Part 1: Demo of Nested ANOVAs

## Load libraries
```{r, message=FALSE}
library(psych) # using for describe() function
library(tidyverse) # using for read_csv() function, so character variables still kept as character
```

## Read data
```{r, message=FALSE}
nested_demo<-read_csv(file="Nested_demo.csv")
```

## Check how variables are read
```{r}
str(nested_demo)
```

## Get variable descriptives
```{r, warning = FALSE}
describe(nested_demo) #note: we get NaN because variables are character (data includes letters)
```


## Calculate the mean values of Y by hospital, drug, and condition
```{r}
# aggregate function here just shwows the values that we're comparing, but not saving these values
aggregate(nested_demo$Y, by=list(Hospital=nested_demo$Hospital), mean)
aggregate(nested_demo$Y, by=list(Drug=nested_demo$Drug), mean)
aggregate(nested_demo$Y, by=list(Condition=nested_demo$Condition), mean)
```

## (5)Nested ANOVA in which Y is regressed on Hospital, Drug, Condition, and the
## (5a) Create ANOVA table that lists all of the factors and the df for each

Factors (i.e., effects we can test with this model)              
Hospital                
Drug                    
Condition within Drug 
Hospital * Drug
Hospital * (Condition within Drug)
Error

## (5b) Conduct a Hospital x Drug ANOVA, ignoring Condition
```{r}
mod1 <- anova(lm(Y~Hospital + Drug + Hospital*Drug, data=nested_demo))
mod1
```

## 5c Conduct separate ANOVA for effect of condition
```{r}
mod2 <- anova(lm(Y~Drug/Condition + Hospital*(Drug/Condition), data=nested_demo))
mod2 # get same model output as for the full model
```


## Full model code (Not doing in demo though)
```{r}
mod_full <- anova(lm(Y~Hospital + Drug + Drug/Condition + Hospital*Drug + Hospital*(Drug/Condition), data=nested_demo))
mod_full

# Syntax: / specifies within, and * specifies the interaction
# Drug not nested within hospital in this data, because the drugs are the same between the two hospitals (i.e., the effects of a particular drug could be different from one hospital to the next).

# Similarly, condition is not nested within hospital in this data, because the conditions are the same between the two hospitals (i.e., the effects of a particular condition could be different from one hospital to the next).

# However, condition IS nested within drug, becuause the conditions are different between the two drugs. Since each condition was not tested for each drug, it is not possible to assess the possibility that the different conditions are different between the drugs.
```

# Part 2: Try it Yourself - NEEDS REVISION
## Read data
```{r, message=FALSE}
nested_practice<-read_csv(file="Nested_practice.csv")
```
 
## Check how variables are read
```{r}
str(nested_practice)
```
 
 
## Get variable descriptives
```{r, row.print=13, warning = FALSE}
psych::describe(nested_practice) #note: variables are character (include text)
```

## Nested ANOVA in which Y is regressed on hospital, drug, and condition, while taking into account nesting structure
```{r}
anova(lm(Y ~ Hospital + Hospital/Drug + Hospital/Condition +  Hospital/Drug/Condition, data=nested_practice))

# drug is nested within hospital, condition is nested within drug, and both are nested within hospital
```

