---
title: "PSY 653 Module 3: Nested Designs"
subtitle: "Gemma Wallace & Neil Yetz"
output: html_notebook
---

# Part 1: Demo of Nested ANOVAs

## Load libraries
```{r, message=FALSE}
library(psych)
library(tidyverse)
```

## Read data
```{r, message=FALSE}
nested_demo<-read.csv(file="Nested_demo.csv")
```
 
## Get variable descriptives
```{r, row.print=13}
psych::describe(nested_demo) #note: variables are factors (include text)
```

# Calculate the mean values of Y by hospital, drug, and condition
```{r}
# aggregate function here just shwows the values that we're comparing, but not saving these values
aggregate(nested_demo$Y, by=list(Hospital=nested_demo$Hospital), mean)
aggregate(nested_demo$Y, by=list(Drug=nested_demo$Drug), mean)
aggregate(nested_demo$Y, by=list(Condition=nested_demo$Condition), mean)
```

## Nested ANOVA in which Y is regressed on Hospital, Drug, Condition, and the interaction between hospital and condition (taking into )
```{r}
# Syntax: / specifies within, and * specifies the interaction
anova(lm(Y ~ Hospital + Drug + Hospital*Drug +  Condition/Drug + Hospital *(Condition/Drug), data=nested_demo))

# drug not nested within hospital in this data, because the drugs are the same between the two hospitals
```

# Part 2: Try it Yourself
## Read data
```{r, message=FALSE}
nested_practice<-read.csv(file="Nested_practice.csv")
```
 
## Get variable descriptives
```{r, row.print=13}
psych::describe(nested_practice) #note: variables are factors (include text)
```

## Nested ANOVA in which Y is regressed on hospital, drug, and condition, while taking into account nesting structure
```{r}
anova(lm(Y ~ Hospital + Hospital/Drug + Hospital/Condition +  Hospital/Drug/Condition, data=nested_practice))

# drug is nested within hospital, condition is nested within drug, and both are nested within hospital
```

